Towards a standardized evaluation of multiple imputation routines

### 2022-02-03 Manuscript/StandardizedEvaluation.Rmd
Add CI and RMSE for all methods in sim

@@

In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng1994multiple. The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR). Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable see e.g. @rubin1976inference, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.

@@

```{r sim_mech, echo=FALSE, cache=TRUE, paged.print=TRUE}
# setup
set.seed(1)
n_sim <- 10
correlations <- c(0, .4, .8)
mechanisms <- c("MCAR", "MAR", "MNAR")

@@

First, developers often only inquire about the 'accuracy' (i.e. how well can the method reproduce the original data). The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. We are interested in the correct answer to the research question; not in the truth itself. This means that, given the framework provided by @rubin1987, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied. After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neyman1934two, p. 591.

@@

Third, sampling variation has always been an essential part of the evaluation of multiple imputation methodology. However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink2014pooling. After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.

@@

Next, missing data should be induced conform a model that is dependent on the observed data. A straightforward technique for inducing different forms of univariate MAR missingness is described in @fimd, p. 63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR missingness can be found in @van2006fully, Appendix B, and @brand1999development, par. 5.2.3. If the missingness is to be induced in longitudinal data, auto-regressive MAR models e.g. cf. @shara2015randomly, model 2 and model 3, can be useful.

>>>>

```{r sim_nobs, echo=FALSE, cache=TRUE, paged.print=TRUE}
# setup
set.seed(1)
n_sim <- 2

sim_nobs <- purrr::map_dfr(1:n_sim, function(.i){
  purrr::map_dfr(c(50, 500, 5000), ~{
    run(n_obs = .x) %>% cbind(sim = .i, .)})})

gtsummary::tbl_summary(sim_nobs, by = method)
sim_mech %>% 
  group_by(mis_mech, corr) %>% 
  summarise(
    mean_cca = mean(mean_cca - 10),
    mean_est = mean(mean_est - 10),
    beta_cca = mean(beta_cca - corr),
    beta_est = mean(beta_est - corr),
    beta_ciw = mean(beta_ciw),
    beta_cr = mean(beta_cr),
    rmse_pred = mean(rmse_pred),
    rmse_cell = mean(rmse_cell))
```

@@

In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng94. The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR). Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable see e.g. @rubi76, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.

@@

```{r sim_mech_corr, echo=FALSE, cache=TRUE, paged.print=TRUE}
# setup
set.seed(1)
n_sim <- 10
correlations <- c(0, .4, .8)
mechanisms <- c("MCAR", "MAR", "MNAR")

@@

First, developers often only inquire about the 'accuracy' (i.e. how well can the method reproduce the original data). The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. We are interested in the correct answer to the research question; not in the truth itself. This means that, given the framework provided by @rubi87, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied. After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neym34, p. 591.

@@

Third, sampling variation has always been an essential part of the evaluation of multiple imputation methodology. However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink14. After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.

@@

Next, missing data should be induced conform a model that is dependent on the observed data. A straightforward technique for inducing different forms of univariate MAR missingness is described in @fimd, p. 63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR missingness can be found in @buur06, Appendix B, and @brand1999development, par. 5.2.3. If the missingness is to be induced in longitudinal data, auto-regressive MAR models e.g. cf. @shara2015randomly, model 2 and model 3, can be useful.
